\chapter{Background and Literature Review} \label{sec:BackgroundAndLiteratureReview}

\section{Background} \label{sec:Background}

Many stories within video games are experienced linearly, with the player
having very little autonomy over how the setting progresses. Some games offer you
set choices throughout the playthrough and\slash or multiple endings, but still ultimately
offer a limited set of experiences. Many players value the replayability of a game,
referring to how much the game can be replayed and explored before becoming stale,
which is known to be difficult to combine with a strong narrative experience.

One way replayability can be introduced is through \textbf{emergent narratives}. These
can take many forms and are frequently a side effect of complex game mechanics. For
example, in 2005, an unintended interaction within World of Warcraft, a game where
thousands of players interact within an online multiplayer world, caused a plague
to spread and started a pandemic. \cite{andrei_weird_2020} None of this behavior was
written into the game as an intended storyline. Rather, it emerged from the individual
behaviors of players, creating a sequence of events that would never have existed
otherwise.

It is also possible to target emergent narratives as a core part of your game.
By shifting focus away from the story as a whole and onto the characters themselves,
much of the complexity that comes from branching storylines can, in theory, be avoided. Instead
of considering every character over every narrative path, the characters can be written
to create each path on an individual level, freeing up writer time.
Significant progress has been made towards being able to represent characters in this
manner, but developers still lack a unified tool that can support this system.

Many tools proposed contain some sort of \textbf{drama management}, which oversees the
overall tension of the story. For example, if events become too dull, the drama
manager might insert some sort of unexpected occurrence to make the plot engaging
again. In extreme cases, the drama manager may be in charge of the rising tension,
climax, and resolution of the plot entirely. This can be thought of as an
all-powerful entity that ensures the narrative stays controlled, to whatever
extent the developer would like.

Emergent narrative is a spectrum in many ways, and development and research exist
on all parts of it. This project aims to create a system where characters act in
highly individualized ways, with the player able to step in for anyone and influence
the story however they'd like. During preliminary research, we went through
existing projects to determine which features were important for our implementation
and what we should avoid during development.

\section{Primary Inspirations} \label{sec:PrimaryInspirations}

There are two primary projects that pioneered the technique we are attempting to
model here: Comme il Faut (2011) and Versu (2012). Both, and subsequent follow-up
projects, were referred to heavily while designing Praxsmith.

Comme il Faut (2011), also referred to as CiF and used to create Prom Week \cite{noauthor_prom_nodate},
saw characters following certain predefined social exchanges to create a social simulation,
referring to the technique as "social physics". \cite{mccoy_comme_2011} This
implementation was not available to the public, but it did provide some vital progress
in the field. Comme il Faut uses a set of databases to store information about the
world state and introduces the idea of social exchanges, with required conditions
determined by predicate logic and a set of effects when a particular branch is chosen.
It is also able to keep track of past events and refer to them, weighting certain
actions based on current and past world states. Although powerful, Comme il Faut
did have some restrictive limitations. For example, social exchanges were limited to
three characters and provided fairly rigid bounds of operation. It was noted by a
future work that rumor spreading was made impossible by this restriction, which would
have been a welcome addition to the simulation. \cite{samuel_ensemble_2015}

Versu (2012) was another private project that attempted to create believable au-
tonomous characters. It involved a lower-level logic language called Praxis, as well
as a higher-level writer-friendly language named Prompter. \cite{evans_ai_2015}
\cite{nelson_prompter_2014} The results from this project seem promising, but are
largely lost to time. The tree and exclusion logic inspired world are powerful,
but they suffer from a lack of structure that could make it more susceptible to bugs.
Versu also introduces the idea of character-driven goals, rather than weights
within social exchanges, as Comme il Faut was designed. These goals can be used in
conjunction with a planner, allowing for characters to make informed decisions about
the actions available to them. This is accomplished through a transactional world state that can
be rolled back after an action is evaluated, allowing characters to ``see into the future.''
Most of the syntax and patterns within this project are pulled from Versu's Praxis,
including sentence syntax, goals, and relation types.

The first social simulation engine available to the public was the Ensemble Engine,
created in 2015. \cite{samuel_ensemble_2015} It fixes a number of the flexibility
issues Comme il Faut had, such as allowing rulesets to take on any number of
participants and allowing the author to create unique categories of social facts.
It also provides comprehensive authoring and debugging user interfaces that allow
authors to wrap their head around complex social simulation mechanics. Designed by
an experienced Comme il Faut developer, these interfaces provide useful references
and show that tooling is crucial for authors. Unfortunately, the Ensemble Engine
project hasn’t received support since 2019.

Praxish was a recent attempt to recreate the Praxis language, but had a gap in
maintenance during the development of this project after its creation in 2023.
It largely maintains the same issues that Praxis had, while not being as fully featured.
\cite{dameris_praxish_2023} Inspired by this project, implementing Praxis in a
modern setting, with additions from other social simulation projects, will be the
primary focus of this research.

\section{Other Related Work} \label{sec:OtherRelatedWork}

Façade, created in 2003, is often considered one of the earlier attempts at creating an
emergent narrative. Instead of focusing heavily on the characters themselves, Façade
organizes its story into beats, which are pulled by a central drama manager. While
the results are impressive, they are not perfect. Two man-years of effort only resulted
in a 20-minute one-act play replayable 6 or 7 times before it was exhausted. The text
input was also notoriously finicky, largely a result of the year it was
created. \cite{mateas_facade_2003}

Some other related works include Crosston Tavern (2021) which aims to let characters
take independent action according to goals. \cite{oliver_crosston_2021} Ryan et al. (2015) explore a
system that allows characters to misremember and lie, allowing for even more dramatic
moments. \cite{ryan_toward_2015} This is an interesting extension, but it is outside the scope of
this project. Inform 7, created in 2006, is a popular open-source language for writing
interactive fiction and was a large inspiration for Versu’s higher-level Prompter
language. \cite{inform_7_inform_nodate} ”ink”, created in 2015, is another such system, allowing for narrative
scripting within games using a higher-level writer-friendly language, also supporting
popular game engine integration. \cite{noauthor_ink_nodate} The Sims is a popular video game series that
centers around individual non-player-controlled characters, with the first installment
released in 2000. \cite{arts_sims_2016} Although largely successful, The Sims does not attempt to
create an overarching personal narrative with the characters at the center and is more focused on
simulation-style gameplay.
