\chapter{Conclusion} \label{sec:Conclusion}

\section{Contributions} \label{sec:Contributions}

This paper provides an overview of what technologies and principles outline
a successful character-narrative DSL. It also creates a new framework, Praxsmith, that
aims to realize these concepts in a modern setting, including:

\begin{itemize}
    \item A flexible game engine built on top of a Rust backend and a web app frontend.
    \item An actor-centric DSL inspired by Praxis, as well as many other influences
    outlined in \Cref{sec:BackgroundAndLiteratureReview}.
    \item Modern tooling, such as build-time type-checking to prevent runtime errors
    and minimize logic bugs.
    \item An authoring environment built into the web app.
    \item Thorough documentation about the features and design decisions of the
    language.
    \item An evaluation of usability and capability, comprising of an author evaluation
    plus a limited user study.
\end{itemize}

\section{Research Question Answers} \label{sec:ResearchQuestionAnswers}

\textbf{What elements make up a successful character-focused narrative DSL?}

As explored in \Cref{sec:PrimaryInspirations}, there are many contributors to a
successful character-focused narrative DSL.

\begin{itemize}
    \item \textbf{Low Verbosity:} Praxsmith suffers from a bloating issue, where
    even simple practices can end up taking a lot of space and being hard to parse
    through. Both Comme il Faut and Versu tackle this problem
    in unique ways with great success from an authoring standpoint.
    \item \textbf{Catching Errors:} Praxsmith has strong typing systems, which
    noticeably catch many mistakes whilst developing a game. The Ensemble Engine
    addresses this by having a debugger that was designed by a core developer
    after noticing a strong need for it.
    \item \textbf{Moderate Flexibility:} Comme il Faut, as noted by the Ensemble
    Engine authors, ended up being overly restrictive in the social simulation
    space. On the other end, Praxis contains some freedoms that could easily lead
    to runtime errors or logic bugs. This project attempts to balance between the
    two, but more research should be done into how effective the approaches it
    takes are.
\end{itemize}

\textbf{Is a structured relationship graph a useful way to model the world state for
writing autonomous characters?}

Initial evidence indicates that this way of modeling the world is effective.
Both Comme il Faut and Versu model their worlds in a similar way, although
not explicitly. Through the usage of Praxsmith, it became clear to both the survey
participants that considering the social space as a network of
relations between agents is an intuitive way to consider the world and opens up
the door to powerful possibilities.

\textbf{Can the fundamentals of Praxis and Comme il Faut be created
in a modernized, type-safe scripting language?}

Although Praxsmith lacked some crucial features, its fundamental construction proves
that it is possible to modernize these concepts. This includes features like the concretely defined
yet flexible relation types and the strict type checker, plus the layered nature of the
project itself, allowing for integration with other engines.

\textbf{Is a logic-level DSL like Praxis usable for creating convincing
emergent narrative simulations?}

Through our experimentation and the user study, it appears that although Praxsmith is
technically capable of complex simulation, it lacks the convenience required by these
scenarios. Authoring quickly becomes hard to manage with even smaller scenarios, so
in situations where there are thousands of relations governing the state of the world,
it would likely become unreasonable to work with. As Comme il Faut and Versu both have
more authorable counterparts, it appears as though a higher-level layer would be required
for Praxsmith to be a truly applicable system.

\section{Limitations} \label{sec:Limitations}

\subsection{Incomplete Survey} \label{sec:IncompleteSurvey}

Due to timeline constraints, the survey and experiment trials conducted for this
project provided incomplete results. Only 4 people completed the trials all the
way through, and the average time spent on the project was around 1 hour, despite
a target of 3 to 6 hours.

With an involved project such as a development environment\slash game engine, it
is difficult to find active participants willing to put in the time necessary to
become familiar with a new technology. Further research could offer an incentive,
such as monetary rewards, to drive longer participation times. Alternatively, a
larger-scale game project could be developed alongside the engine, which would allow for
direct feedback from the authors most involved in the process. Larger projects
like Comme il Faut \cite{mccoy_social_2014} and Versu \cite{nelson_prompter_2014} were
also developed this way, with ``Prom Week'' and ``Blood and Laurels'' being
the associated game projects, respectively.

\subsection{Survey Bias} \label{sec:SurveyBias}

This survey came with a significant amount of selection bias, as all respondents
to the preliminary survey had high experience with programming languages,
likely due to the areas in which this project was advertised. An important part of
this project was how easy the system is to use for writers, which ended up not
being determinable from the selection of survey participants. For future research,
more trials should be run specifically with writers who are not familiar
with programming for a more holistic interpretation.

The user response survey was a limiting factor on this research project. As stated
in \Cref{sec:SurveyBias}, the outreach methods of the survey caused it to be skewed
towards those who were already comfortable with programming, which means the writers
who would be using the tool were not evaluated.

\subsection{Personal Bias} \label{sec:PersonalBias}

As multiple of the trial participants were acquainted with us to varying
degrees, there is personal bias embedded within the results as well. Casual channels
of communication (Discord messages) were used during the experimentation stage for
questions and feedback about the systems. Although all participants completed the
surveys as intended, this personal bias cannot be overlooked and should be addressed
in future research by conducting a more repeatable experiment.

\section{Future Work} \label{sec:FutureWork}

\subsection{Improved Studies} \label{sec:ImprovedStudies}

To address the survey limitations enumerated in \Cref{sec:Limitations}, more thoroughly
conducted studies could be used. By offering an incentive to participants, broadening the
pool of outreach, and allowing for more time to experiment, survey results would cover
a much more accurate representation of the intended audience.

\subsection{Long-Term Project} \label{sec:LongTermProject}

Another avenue to resolve the evaluation limitations is to trial a longer-term project
using Praxsmith to evaluate how it works in a real-world scenario. As noted in \Cref{sec:Limitations},
Comme il Faut \cite{mccoy_social_2014} and Versu \cite{nelson_prompter_2014} had similar
associated projects, allowing for the developers to build off of their own experiences
as users of the system and more accurately evaluate its capabilities and weaknesses.

\subsection{Engine And Interface Improvements} \label{sec:EngineAndInterfaceImprovements}

There are also many potential improvements to the engine and language that would
make it easier to use from the developer standpoint, as outlined in \Cref{sec:Evaluation}.

Perhaps the foremost technical opportunity for future research is expanding upon the
authorability of Praxsmith by adding either:

\begin{itemize}
    \item A higher-level language that writes like a play script, something
    authors are likely familiar with. This is what Prompter does for the Versu engine.
    \cite{nelson_prompter_2014} This would likely give much better results for a
    survey of writers working with this project, and it would still give developers
    the flexibility to realize the full potential of the engine.
    \item An authoring user interface that addresses some of the verbosity and
    bloating observed during usage of the Praxsmith language. The authors of the
    Ensemble Engine recognized this limitation and addressed it with a fully featured
    interface, \cite{samuel_ensemble_2015} so a similar extension is worth looking
    into for Praxsmith, potentially replacing the scripting language altogether.
\end{itemize}
